---
title: "Final Analysis"
output: pdf_document
date: "April 5"
---

```{r}
library(here)
library(lubridate)
library(tidyverse)
library(dplyr)
library(ggplot2); library(patchwork)
library(vcdExtra)
library(moments) 

#importing csvs
winter_ozone_egrid_all_df <- read.csv('winter_ozone_egrid_all.csv')
summer_ozone_egrid_all_df <- read.csv('summer_ozone_egrid_all.csv')

winter_PM25_egrid_all_df <- read.csv('winter_PM25_egrid_all.csv')
summer_PM25_egrid_all_df <- read.csv('summer_PM25_egrid_all.csv')

#dont need anymore i think

# winter_ozone_egrid_all_df <- winter_ozone_egrid_all_df %>%
#   #filter(Plant.state.abbreviation %in% c('GA','FL','NC','SC','MS','TN','AL')) %>%
#   #to get rid of negative numbers and 0s ...
#   mutate(Sum_Plant_NetGenMW = ifelse(Sum_Plant_NetGenMW <= 0, NA, Sum_Plant_NetGenMW))
#  
# summer_ozone_egrid_all_df <- summer_ozone_egrid_all_df %>%
#   #filter(Plant.state.abbreviation %in% c('GA','FL','NC','SC','MS','TN','AL')) %>%
#   mutate(Sum_Plant_NetGenMW = ifelse(Sum_Plant_NetGenMW <= 0, NA, Sum_Plant_NetGenMW))
# 
# winter_PM25_egrid_all_df <- winter_PM25_egrid_all_df %>%
#   filter(Plant.state.abbreviation %in% c('GA','FL','NC','SC','MS','TN','AL')) %>%
#   mutate(Sum_Plant_NetGenMW = ifelse(Sum_Plant_NetGenMW <= 0, NA, Sum_Plant_NetGenMW))
# 
# summer_PM25_egrid_all_df <- summer_PM25_egrid_all_df %>%
#   filter(Plant.state.abbreviation %in% c('GA','FL','NC','SC','MS','TN','AL')) %>%
#   mutate(Sum_Plant_NetGenMW = ifelse(Sum_Plant_NetGenMW <= 0, NA, Sum_Plant_NetGenMW))

```

# Ozone Wrangling
```{r}
winter_ozone <- winter_ozone_egrid_all_df %>%
  mutate(Season = 'Winter') %>%
  mutate(SeasonBinary = '0')

summer_ozone <- summer_ozone_egrid_all_df %>%
  mutate(Season = 'Summer')%>%
  mutate(SeasonBinary = '1')
  
all_ozone <- bind_rows(summer_ozone, winter_ozone)

```

# PM 2.5 Wrangling
```{r}
# NOTE - PM winter and summer have same amount of data points in winter and summer - which could be better for analysis?
# while ozone is a bit much
winter_PM25 <- winter_PM25_egrid_all_df %>%
  mutate(Season = 'Winter') %>%
  mutate(SeasonBinary = '0')

summer_PM25 <- summer_PM25_egrid_all_df %>%
  mutate(Season = 'Summer')%>%
  mutate(SeasonBinary = '1')
  
all_PM25 <- bind_rows(summer_PM25, winter_PM25)

```

# Log transforming data
```{r}
all_PM25 <- all_PM25 %>%
  mutate(LogSum_Plant_NetGenMW = log10(Sum_Plant_NetGenMW))

all_ozone <- all_ozone %>%
  mutate(LogSum_Plant_NetGenMW = log10(Sum_Plant_NetGenMW))
```

some limitations... i dont think we should be using annual generation because that is the net, so its minusing the gross - net so we arent getting TOTAL electricity generation that plants is consuming thats why some are so low and some are negative... so we arent using just total generation these plants are producing :---)
so we need to write in paper about this probably?


# Exploratory graphs 
## Ozone v. Net Gen - Trends with Seasonlity (2)
```{r}
# OZONE
#by mean ozone concentration on y and net gen on x
all_ozone_fig1<- all_ozone %>%
  ggplot(aes(x = Sum_Plant_NetGenMW, y = average_month_ozone_concentration, color = Season)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  scale_x_continuous(labels = scales::comma) #no scientific notation +
  #scale_x_log10() +
  labs(x = 'Mean Net Generation of all Counties (MW)', y = 'Average Ozone Concentrations (unit)',
       title = 'Average Net Generation and \n Ozone Concentrations by County in Summer and Winter')
all_ozone_fig1

#log10 ed 
all_ozone_fig1scaled<- all_ozone %>%
  ggplot(aes(x = Sum_Plant_NetGenMW, y = average_month_ozone_concentration, color = Season)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  scale_x_log10() +
  labs(x = 'Mean Net Generation of all Counties (MW)', y = 'Average Ozone Concentrations (unit)',
       title = 'Average Net Generation and \n Ozone Concentrations by County in Summer and Winter')
all_ozone_fig1scaled


```

## PM 2.5 v. Net Gen - Trends with Seasonality (2)
```{r}
# PM 2.5 
#by mean PM 2.5 concentration on y and net gen on x
all_PM25_fig1<- all_PM25 %>%
  ggplot(aes(x = Sum_Plant_NetGenMW, y = average_month_PM25_concentration, color = Season)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  #scale_x_log10() +
  scale_x_continuous(labels = scales::comma) #no scientific notation +
  labs(x = 'Mean Net Generation of All Counties (MW)', y = 'Average PM 2.5 Concentrations (unit)',
       title = 'Average Net Generation and \n PM 2.5 Concentrations by County in Summer and Winter')
all_PM25_fig1

# log 10 ed
all_PM25_fig1scaled<- all_PM25 %>%
  ggplot(aes(x = Sum_Plant_NetGenMW, y = average_month_PM25_concentration, color = Season)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  scale_x_log10() +
  labs(x = 'Mean Net Generation of All Counties (MW)', y = 'Average PM 2.5 Concentrations (unit)',
       title = 'Average Net Generation and \n PM 2.5 Concentrations by County in Summer and Winter')
all_PM25_fig1scaled
```


# Histograms - PM / Ozone Count with Seasonality (2)
maybe perform log transformation - really just looking at ozone/pm with seasonality
```{r}
# ozone histogram
ozone_histogram <- ggplot(all_ozone, aes(x = average_month_ozone_concentration, 
                                        fill = Season)) + 
  geom_histogram() + 
  scale_fill_manual(values = c('orange',"blue")) + 
  labs(x = 'Ozone Concentrations', y = "Count",
       title = 'Ozone Concentrations by Month') + 
  facet_grid(.~Season) + # facets by habitat to avoid overlap 
  theme_bw() + # removes the grey background 
  theme(legend.position = "none") #removes legend

ozone_histogram
```

```{r}
# PM 2.5 histogram
PM25_histogram <- ggplot(all_PM25, aes(x = average_month_PM25_concentration, 
                                        fill = Season)) + 
  geom_histogram() + 
  scale_fill_manual(values = c('orange',"blue")) + 
  labs(x = 'PM 2.5 Concentrations', y = "Count",
       title = 'PM 2.5 Concentrations by Month') + 
  facet_grid(.~Season) + # facets by habitat to avoid overlap 
  theme_bw() + # removes the grey background 
  theme(legend.position = "none") #removes legend
PM25_histogram
```

Tidy Census histograms?
```{r}

```


# Skewness and Kurtosis for Ozone and PM - Seasonlity (4)
```{r}
# Ozone 
ozonewinter_data <- all_ozone %>% filter(Season == "Winter")
skewness(ozonewinter_data$average_month_ozone_concentration, na.rm = TRUE) #-0.8506
kurtosis(ozonewinter_data$average_month_ozone_concentration, na.rm = TRUE) #3.5699

ozonesummer_data <- all_ozone %>% filter(Season == "Summer")
skewness(ozonesummer_data$average_month_ozone_concentration, na.rm = TRUE) #0.175
kurtosis(ozonesummer_data$average_month_ozone_concentration, na.rm = TRUE) #2.3717
```

```{r}
# PM 2.5
pm25winter_data <- all_PM25 %>% filter(Season == "Winter")
skewness(pm25winter_data$average_month_PM25_concentration, na.rm = TRUE) #-1.165
kurtosis(pm25winter_data$average_month_PM25_concentration, na.rm = TRUE) #5.4427

pm25summer_data <- all_PM25 %>% filter(Season == "Summer")
skewness(pm25summer_data$average_month_PM25_concentration, na.rm = TRUE) #0.182
kurtosis(pm25summer_data$average_month_PM25_concentration, na.rm = TRUE) #2.708
```


# Linear Regression Beginning

## Define Research Questions

## Examine Data
```{r}
# Create and examine histograms of raw values.
(hist1_netgen <- ggplot(all_ozone, aes(x = Sum_Plant_NetGenMW)) 
 + geom_histogram() +
   theme_bw())
  #SUPERRR skewed - should handle this - log transform?

(hist1_netgenlogscale <- ggplot(all_ozone, aes(x = LogSum_Plant_NetGenMW)) + geom_histogram() +
   theme_bw())

(hist2_ozoneconcent <- ggplot(all_ozone, aes(x = average_month_ozone_concentration)) + 
    geom_histogram() +
    theme_bw())

(hist3_pm25concent <- ggplot(all_PM25, aes(x = average_month_PM25_concentration)) + 
    geom_histogram() +
    theme_bw())

# Create and examine an initial scatterplot of data.
(scatter1_ozone <- ggplot(all_ozone, aes(x = LogSum_Plant_NetGenMW, y = average_month_ozone_concentration)) +
    geom_point() + 
    theme_bw())

(scatter2_pm25 <- ggplot(all_PM25, aes(x = LogSum_Plant_NetGenMW, y = average_month_PM25_concentration))
  +
    geom_point() + 
    theme_bw())

```


```{r}

```

