---
title: "Final Analysis"
output: pdf_document
date: "April 5"
---

```{r}
library(here)
library(lubridate)
library(tidyverse)
library(dplyr)
library(ggplot2); library(patchwork)

winter_ozone_egrid_all_df <- read.csv('winter_ozone_egrid_all.csv')
summer_ozone_egrid_all_df <- read.csv('summer_ozone_egrid_all.csv')

winter_PM25_egrid_all_df <- read.csv('winter_PM25_egrid_all.csv')
summer_PM25_egrid_all_df <- read.csv('summer_PM25_egrid_all.csv')

#dont need anymore i think

# winter_ozone_egrid_all_df <- winter_ozone_egrid_all_df %>%
#   #filter(Plant.state.abbreviation %in% c('GA','FL','NC','SC','MS','TN','AL')) %>%
#   #to get rid of negative numbers and 0s ...
#   mutate(Sum_Plant_NetGenMW = ifelse(Sum_Plant_NetGenMW <= 0, NA, Sum_Plant_NetGenMW))
#  
# summer_ozone_egrid_all_df <- summer_ozone_egrid_all_df %>%
#   #filter(Plant.state.abbreviation %in% c('GA','FL','NC','SC','MS','TN','AL')) %>%
#   mutate(Sum_Plant_NetGenMW = ifelse(Sum_Plant_NetGenMW <= 0, NA, Sum_Plant_NetGenMW))
# 
# winter_PM25_egrid_all_df <- winter_PM25_egrid_all_df %>%
#   filter(Plant.state.abbreviation %in% c('GA','FL','NC','SC','MS','TN','AL')) %>%
#   mutate(Sum_Plant_NetGenMW = ifelse(Sum_Plant_NetGenMW <= 0, NA, Sum_Plant_NetGenMW))
# 
# summer_PM25_egrid_all_df <- summer_PM25_egrid_all_df %>%
#   filter(Plant.state.abbreviation %in% c('GA','FL','NC','SC','MS','TN','AL')) %>%
#   mutate(Sum_Plant_NetGenMW = ifelse(Sum_Plant_NetGenMW <= 0, NA, Sum_Plant_NetGenMW))

```

# Ozone
```{r}
winter_ozone <- winter_ozone_egrid_all_df %>%
  mutate(Season = 'Winter') %>%
  mutate(SeasonBinary = '0')

summer_ozone <- summer_ozone_egrid_all_df %>%
  mutate(Season = 'Summer')%>%
  mutate(SeasonBinary = '1')
  
all_ozone <- bind_rows(summer_ozone, winter_ozone)

```

# PM 2.5
```{r}
# NOTE - PM winter and summer have same amount of data points in winter and summer - which could be better for analysis?
# while ozone is a bit much
winter_PM25 <- winter_PM25_egrid_all_df %>%
  mutate(Season = 'Winter') %>%
  mutate(SeasonBinary = '0')

summer_PM25 <- summer_PM25_egrid_all_df %>%
  mutate(Season = 'Summer')%>%
  mutate(SeasonBinary = '1')
  
all_PM25 <- bind_rows(summer_PM25, winter_PM25)

```

some limitations... i dont think we should be using annual generation because that is the net, so its minusing the gross - net so we arent getting TOTAL electricity generation that plants is consuming thats why some are so low and some are negative... so we arent using just total generation these plants are producing :---)
so we need to write in paper about this probably?


# Exploratory graphs 
## Ozone 
```{r}
# OZONE
#by mean ozone concentration on y and net gen on x
all_ozone_fig1<- all_ozone %>%
  ggplot(aes(x = Sum_Plant_NetGenMW, y = average_month_ozone_concentration, color = Season)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  scale_x_continuous(labels = scales::comma) #no scientific notation +
  #scale_x_log10() +
  labs(x = 'Mean Net Generation of all Counties (MW)', y = 'Average Ozone Concentrations (unit)',
       title = 'Average Net Generation and \n Ozone Concentrations by County in Summer and Winter')
all_ozone_fig1


all_ozone_fig1scaled<- all_ozone %>%
  ggplot(aes(x = Sum_Plant_NetGenMW, y = average_month_ozone_concentration, color = Season)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  scale_x_log10() +
  labs(x = 'Mean Net Generation of all Counties (MW)', y = 'Average Ozone Concentrations (unit)',
       title = 'Average Net Generation and \n Ozone Concentrations by County in Summer and Winter')
all_ozone_fig1scaled


```

## PM 2.5 
```{r}
# PM 2.5 
#by mean PM 2.5 concentration on y and net gen on x
all_PM25_fig1<- all_PM25 %>%
  ggplot(aes(x = Sum_Plant_NetGenMW, y = average_month_PM25_concentration, color = Season)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  #scale_x_log10() +
  scale_x_continuous(labels = scales::comma) #no scientific notation +
  labs(x = 'Mean Net Generation of All Counties (MW)', y = 'Average PM 2.5 Concentrations (unit)',
       title = 'Average Net Generation and \n PM 2.5 Concentrations by County in Summer and Winter')
all_PM25_fig1


all_PM25_fig1scaled<- all_PM25 %>%
  ggplot(aes(x = Sum_Plant_NetGenMW, y = average_month_PM25_concentration, color = Season)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  scale_x_log10() +
  labs(x = 'Mean Net Generation of All Counties (MW)', y = 'Average PM 2.5 Concentrations (unit)',
       title = 'Average Net Generation and \n PM 2.5 Concentrations by County in Summer and Winter')
all_PM25_fig1scaled
```


