---
title: "multi_lvl_models"
output: html_document
---

```{r}
library(here)
library(tidyverse)
library(lubridate)
library(nlme)
library(gtsummary)
```

Notes:

Log transform predictors:
- Annual generation
- Population

Doesn't need log transformation (didn't see much shift in scatter + lm exploratory graphs):
- Percent no vehicle
- Median Income
- Housing burden

PM25_lvl ~ annual generation + population + median income + percent of household with no cars + percent households spending more than 30% of income on housing costs + winter + summer , random = ~1 | State, data = all_of_ data

```{r}
#importing csvs
all_ozone_df <- read.csv('ALL_OZONE_FINAL.csv')
all_pm25_df <- read.csv('ALL_PM25_FINAL.csv')

#Making Season column factors
all_ozone_df <- all_ozone_df %>% 
  mutate(Season = factor(Season, levels = c("Winter", "Summer")))

all_pm25_df <- all_pm25_df %>% 
  mutate(Season = factor(Season, levels = c("Winter", "Summer")))

```

#OZONE: Fit linear regression model with random effect & evaluate model diagnostics
```{r}
#Make sure to avoid undefined variables when you do log(0)
all_ozone_df <- all_ozone_df %>%
  mutate(
    LogSum_Plant_NetGenMW = ifelse(Sum_Plant_NetGenMW > 0, log10(Sum_Plant_NetGenMW), NA),
    log_Population = ifelse(Population > 0, log10(Population), NA)
  )

#adding a column in the data to show state
all_ozone_df <- all_ozone_df %>%
  separate(CountyState, into = c("County", "State"), sep = ",\\s*")

ozone_lme1 <- lme(average_month_ozone_concentration ~ Sum_Plant_NetGenMW + 
                  Season + 
                  percent_white + 
                  percent_black + 
                  percent_asian + 
                  percent_housing_burden + 
                  Median_Income + 
                  percent_no_vehicle + 
                  log_Population, random = ~1 | State, 
                  data = all_ozone_df)

#fitting standard linear regression, which allows us to compare if the addition of a random effect yeilds a better model fit
ozone_lm <- gls(average_month_ozone_concentration ~ Sum_Plant_NetGenMW + 
                  Season + 
                  percent_white + 
                  percent_black + 
                  percent_asian + 
                  percent_housing_burden + 
                  Median_Income + 
                  percent_no_vehicle + 
                  log_Population,
                  data = all_ozone_df)

AIC(ozone_lme1,ozone_lm)

summary(ozone_lme1)
plot(ozone_lme1)
qqnorm(ozone_lme1)
```
The multi-level model has a lower AIC, and therefore has the better model fit!

##Communicating the methods and results
```{r}
tbl_regression(ozone_lme1)
```

#PM2.5: Fit linear regression model with random effect & evaluate model diagnostics
```{r}
#Make sure to avoid undefined variables when you do log(0)
all_pm25_df <- all_pm25_df %>%
  mutate(
    LogSum_Plant_NetGenMW = ifelse(Sum_Plant_NetGenMW > 0, log10(Sum_Plant_NetGenMW), NA),
    log_Population = ifelse(Population > 0, log10(Population), NA)
  )

#adding a column in the data to show state
all_pm25_df <- all_pm25_df %>%
  separate(CountyState, into = c("County", "State"), sep = ",\\s*")

pm25_lme1 <- lme(average_month_PM25_concentration ~ Sum_Plant_NetGenMW + 
                  Season + 
                  percent_white + 
                  percent_black + 
                  percent_asian + 
                  percent_housing_burden + 
                  Median_Income + 
                  percent_no_vehicle + 
                  log_Population, random = ~1 | State, 
                  data = all_pm25_df)

#fitting standard linear regression, which allows us to compare if the addition of a random effect yeilds a better model fit
pm25_lm <- gls(average_month_PM25_concentration ~ Sum_Plant_NetGenMW + 
                  Season + 
                  percent_white + 
                  percent_black + 
                  percent_asian + 
                  percent_housing_burden + 
                  Median_Income + 
                  percent_no_vehicle + 
                  log_Population,
                  data = all_pm25_df)

AIC(pm25_lme1,pm25_lm )

summary(pm25_lme1)
plot(pm25_lme1)
qqnorm(pm25_lme1)
```
The multi-level model has a lower AIC, and therefore has the better model fit!

##Communicating the methods and results
```{r}
tbl_regression(pm25_lme1)
```

