---
title: "Stat_Figures"
output: pdf_document
date: "2025-03-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading in main EGrid Data
```{r}
library(here)
library(lubridate)
library(tidyverse)
library(readxl)
library(openxlsx)
library(dplyr)
#reading in egrid 2022 data
egrid_data <- read.xlsx(xlsxFile="./Datafiles/egrid2022_data.xlsx",sheet = "PLNT22")
egrid_data$Plant.FIPS.county.code <- as.numeric(egrid_data$Plant.FIPS.county.code)
egrid_data$`Plant.nameplate.capacity.(MW)` <- as.numeric(egrid_data$`Plant.nameplate.capacity.(MW)`)


#created a dataframe with column name definitions
egrid_names <- egrid_data[1,]
egrid_names <- egrid_names %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(cols = everything() ,names_to = 'Definition', values_to = 'Abbrev') 
  
```
```

# Loading in Ozone Concentration Data for all 7 States
```{r}
#ozone level recordings for all 7 states in 2022
#Florida 
ozone_FL_data <- read.csv('./Datafiles/ozone_FL.csv')
ozone_FL_data$Date <- as.Date(ozone_FL_data$Date, format = "%m/%d/%Y") #converting Date into date
ozone_FL_data <- ozone_FL_data %>% mutate(Month = month(Date)) #making a column with Month number

#Georgia
ozone_GA_data <- read.csv('./Datafiles/ozone_GA.csv')
ozone_GA_data$Date <- as.Date(ozone_GA_data$Date, format = "%m/%d/%Y")
ozone_GA_data <- ozone_GA_data %>% mutate(Month = month(Date))

#Mississippi 
ozone_MS_data <- read.csv('./Datafiles/ozone_MS.csv')
ozone_MS_data$Date <- as.Date(ozone_MS_data$Date, format = "%m/%d/%Y")
ozone_MS_data <- ozone_MS_data %>% mutate(Month = month(Date))

#North Carolina
ozone_NC_data <- read.csv('./Datafiles/ozone_NC.csv')
ozone_NC_data$Date <- as.Date(ozone_NC_data$Date, format = "%m/%d/%Y")
ozone_NC_data <- ozone_NC_data %>% mutate(Month = month(Date))

#South Carolina
ozone_SC_data <- read.csv('./Datafiles/ozone_SC.csv')
ozone_SC_data$Date <- as.Date(ozone_SC_data$Date, format = "%m/%d/%Y")
ozone_SC_data <- ozone_SC_data %>% mutate(Month = month(Date))

#Tennesse
ozone_TN_data <- read.csv('./Datafiles/ozone_TN.csv')
ozone_TN_data$Date <- as.Date(ozone_TN_data$Date, format = "%m/%d/%Y")
ozone_TN_data <- ozone_TN_data %>% mutate(Month = month(Date))

#Alabama
ozone_AL_data <- read.csv('./Datafiles/ozone_AL.csv')
ozone_AL_data$Date <- as.Date(ozone_AL_data$Date, format = "%m/%d/%Y")
ozone_AL_data <- ozone_AL_data %>% mutate(Month = month(Date))

```

# The following are winter and summer for the 7 states
We are doing Winter as January and Summer as July, because in the Southeast they are the coldest and hottest months

Heili notes - #facet grid 
  #scales = fixed vs free
#patchwork () 
 


# Florida 
## FLORIDA WINTER
### Data Wrangling


```{r}
# data wrangling

#Egrid data
egrid_data_FL_counties <- egrid_data %>%
  filter(Plant.state.abbreviation == 'FL') %>% 
  filter(Plantprimar)
  relocate(Plant.FIPS.county.code, 
           Plant.county.name, `Plant.nameplate.capacity.(MW)`) %>% #so i can see what im working with
  select(Plant.FIPS.county.code, 
         Plant.county.name,
         `Plant.nameplate.capacity.(MW)`) %>% #only important ones i think
  rename(County.FIPS.Code = Plant.FIPS.county.code) %>% #to merge egrid & ozone data frames by common col
  group_by(County.FIPS.Code, Plant.county.name) %>% #change group by to new column name
  summarize(Mean_Plant_Nameplate_CapacityMW = mean(`Plant.nameplate.capacity.(MW)`)) %>%
  ungroup()

#Winter ozone data 
winter_ozone_FL_data <-ozone_FL_data %>%
  filter(Month == 1) %>% #january
  group_by(County, County.FIPS.Code) %>%
  summarize(average_month_ozone_concentration = mean(Daily.Max.8.hour.Ozone.Concentration),
            Average_Month_AQI_Value = mean(Daily.AQI.Value)) %>% #not sure which to do/difference rn
  ungroup()


#now combining the two data frames
FL_winter_ozone_egrid <- merge(egrid_data_FL_counties, winter_ozone_FL_data, by = 'County.FIPS.Code')
```

### Graphs
```{r}
#graphs

#by mean ozone concentration on y and nameplate on x
FL_winter_ozone_concen_fig <- FL_winter_ozone_egrid %>%
  ggplot(aes(x = Mean_Plant_Nameplate_CapacityMW, y = average_month_ozone_concentration)) +
   geom_jitter() +
  scale_x_log10() +
  labs(x = 'Mean Nameplate Capacities of Florida Counties (MW)', y = 'Average Ozone Concentrations by Counties (unit)',
       title = 'Florida Average Nameplate Capacities and \n Ozone Concentrations by County in January ')
 FL_winter_ozone_concen_fig
 
 #AQI .. same graph as above but on different time scales on Y .. so may be easier to read AQI ? not sure
 FL_winter_ozone_AQI <- FL_winter_ozone_egrid %>%
  ggplot(aes(x = Mean_Plant_Nameplate_CapacityMW, y = Average_Month_AQI_Value)) +
   geom_jitter() +
   scale_x_log10() + #Heili suggested this!! better for outliers-can set ticks / break points on x axis too
  labs(x = 'Mean Nameplate Capacities of Florida Counties (MW)', y = 'Average AQI (unit)',
       title = 'Florida Average Nameplate Capacities and \n Ozone AQI by County in January ')
 FL_winter_ozone_AQI

```

## FLORIDA SUMMER
### Data Wrangling
```{r}
# data wrangling
#Florida 
summer_ozone_FL_data <-ozone_FL_data %>%
  filter(Month == 7) %>% #july
  group_by(County, County.FIPS.Code) %>%
  summarize(average_month_ozone_concentration = mean(Daily.Max.8.hour.Ozone.Concentration),
            Average_Month_AQI_Value = mean(Daily.AQI.Value)) %>% #not sure which to do/difference rn
  ungroup()

#now combining the two data frames
FL_summer_ozone_egrid <- merge(egrid_data_FL_counties, summer_ozone_FL_data, by = 'County.FIPS.Code')
```

### Graphs
```{r}
#by mean ozone concentration on y and nameplate on x
FL_summer_ozone_concen <- FL_summer_ozone_egrid %>%
  ggplot(aes(x = Mean_Plant_Nameplate_CapacityMW, y = average_month_ozone_concentration)) +
   geom_jitter() +
  scale_x_log10() +
  labs(x = 'Mean Nameplate Capacities of Florida Counties (MW)', y = 'Average Ozone Concentrations by Counties (unit)',
       title = 'Florida Average Nameplate Capacities and \n Ozone Concentrations by County in July ')
 FL_summer_ozone_concen
 
 #AQI .. same graph as above but on different time scales on Y .. so may be easier to read AQI ? not sure
 FL_summer_ozone_AQI <- FL_summer_ozone_egrid %>%
  ggplot(aes(x = Mean_Plant_Nameplate_CapacityMW, y = Average_Month_AQI_Value)) +
   geom_jitter() +
   scale_x_log10() +
  labs(x = 'Mean Nameplate Capacities of Florida Counties (MW)', y = 'Average AQI (unit)',
       title = 'Florida Average Nameplate Capacities and \n Ozone AQI by County in July ')
 FL_summer_ozone_AQI
```


```{r}
#winter continued ...

#Georgia
winter_ozone_GA_data<-ozone_GA_data

#Mississippi 
winter_ozone_MS_data<-ozone_MS_data

#North Carolina
winter_ozone_NC_data<-ozone_NC_data

#South Carolina
winter_ozone_SC_data<-ozone_SC_data

#Tennessee
winter_ozone_TN_data<-ozone_TN_data

#Alabama
winter_ozone_AL_data<-ozone_AL_data

```

# Summer Data (July)




```{r}
#Georgia
summer_ozone_GA_data<-ozone_GA_data

#Mississippi 
summer_ozone_MS_data<-ozone_MS_data

#North Carolina
summer_ozone_NC_data<-ozone_NC_data

#South Carolina
summer_ozone_SC_data<-ozone_SC_data

#Tennesse
summer_ozone_TN_data<-ozone_TN_data

#Alabama
summer_ozone_AL_data<-ozone_AL_data

```

