---
title: "PM_Stat_Figures"
output: pdf_document
date: "2025-04-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading in main EGrid Data
```{r}
library(here)
library(lubridate)
library(tidyverse)
library(readxl); library(openxlsx)
library(dplyr)
library(ggplot2); library(patchwork)

#reading in egrid 2022 data
egrid_data <- read.xlsx(xlsxFile="./Datafiles/egrid2022_data.xlsx",sheet = "PLNT22")
egrid_data$Plant.FIPS.county.code <- as.numeric(egrid_data$Plant.FIPS.county.code)
egrid_data$`Plant.annual.net.generation.(MWh)` <- as.numeric(egrid_data$`Plant.annual.net.generation.(MWh)`)


#created a dataframe with column name definitions
egrid_names <- egrid_data[1,]
egrid_names <- egrid_names %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(cols = everything() ,names_to = 'Definition', values_to = 'Abbrev') 

#filtering
egrid_data <- egrid_data %>%
  filter(!row_number() %in% c(1)) %>% #filtering out first row with names
  filter(Plant.primary.fuel.category %in% c('BIOMASS','GAS','COAL','OIL','OFSL','OTHF'))

```

# Loading in PM Concentration Data for all 7 States
```{r}
#PM level recordings for all 7 states in 2022
#Florida 
PM25_FL_data <- read.csv('./Datafiles/PM25_FL.csv')
PM25_FL_data$Date <- as.Date(PM25_FL_data$Date, format = "%m/%d/%Y") #converting Date into date
PM25_FL_data <- PM25_FL_data %>% mutate(Month = month(Date)) #making a column with Month number

#Georgia
PM25_GA_data <- read.csv('./Datafiles/PM25_GA.csv')
PM25_GA_data$Date <- as.Date(PM25_GA_data$Date, format = "%m/%d/%Y")
PM25_GA_data <- PM25_GA_data %>% mutate(Month = month(Date))

#Mississippi 
PM25_MS_data <- read.csv('./Datafiles/PM25_MS.csv')
PM25_MS_data$Date <- as.Date(PM25_MS_data$Date, format = "%m/%d/%Y")
PM25_MS_data <- PM25_MS_data %>% mutate(Month = month(Date))

#North Carolina
PM25_NC_data <- read.csv('./Datafiles/PM25_NC.csv')
PM25_NC_data$Date <- as.Date(PM25_NC_data$Date, format = "%m/%d/%Y")
PM25_NC_data <- PM25_NC_data %>% mutate(Month = month(Date))

#South Carolina
PM25_SC_data <- read.csv('./Datafiles/PM25_SC.csv')
PM25_SC_data$Date <- as.Date(PM25_SC_data$Date, format = "%m/%d/%Y")
PM25_SC_data <- PM25_SC_data %>% mutate(Month = month(Date))

#Tennesse
PM25_TN_data <- read.csv('./Datafiles/PM25_TN.csv')
PM25_TN_data$Date <- as.Date(PM25_TN_data$Date, format = "%m/%d/%Y")
PM25_TN_data <- PM25_TN_data %>% mutate(Month = month(Date))

#Alabama
PM25_AL_data <- read.csv('./Datafiles/PM25_AL.csv')
PM25_AL_data$Date <- as.Date(PM25_AL_data$Date, format = "%m/%d/%Y")
PM25_AL_data <- PM25_AL_data %>% mutate(Month = month(Date))

#Combine all PM25 data
PM25_data_all <- bind_rows(
  PM25_FL_data, PM25_GA_data, PM25_MS_data, 
  PM25_NC_data, PM25_SC_data, PM25_TN_data, 
  PM25_AL_data
)
```

# The following are winter and summer for the 7 states
We are doing Winter as January and Summer as July, because in the Southeast they are the coldest and hottest months

Heili notes - #facet grid 
  #scales = fixed vs free
#patchwork () 
# Florida 
## FLORIDA WINTER
### Data Wrangling


```{r}
# data wrangling

#Egrid data
egrid_data_FL_counties <- egrid_data %>%
  filter(Plant.state.abbreviation == 'FL') %>% 
  relocate(Plant.FIPS.county.code, 
           Plant.county.name, `Plant.annual.net.generation.(MWh)`) %>% #so i can see what im working with
  select(Plant.FIPS.county.code, 
         Plant.county.name,
         `Plant.annual.net.generation.(MWh)`) %>% #only important ones i think
  rename(County.FIPS.Code = Plant.FIPS.county.code) %>% #to merge egrid & PM25 data frames by common col
  group_by(County.FIPS.Code, Plant.county.name) %>% #change group by to new column name
  summarize(Mean_Plant_NetGenMW = mean(`Plant.annual.net.generation.(MWh)`)) %>%
  ungroup()

#Winter PM25 data 
winter_PM25_FL_data <-PM25_FL_data %>%
  filter(Month == 1) %>% #january
  group_by(County, County.FIPS.Code) %>%
  summarize(average_month_PM25_concentration = mean(Daily.Mean.PM2.5.Concentration),
            Average_Month_AQI_Value = mean(Daily.AQI.Value)) %>% #not sure which to do/difference rn
  ungroup()

#now combining the two data frames
FL_winter_PM25_egrid <- merge(egrid_data_FL_counties, winter_PM25_FL_data, by = 'County.FIPS.Code')
```

### Graphs
```{r}
#graphs

#by mean PM25 concentration on y and net generation on x
FL_winter_PM25_concen_fig <- FL_winter_PM25_egrid %>%
  ggplot(aes(x = Mean_Plant_NetGenMW, y = average_month_PM25_concentration)) +
   geom_jitter() +
  scale_x_log10() +
  labs(x = 'Mean Net Gen Capacities of Florida Counties (MW)', y = 'Average PM25 Concentrations by Counties (unit)',
       title = 'Florida Average Net Gen Capacities and \n PM25 Concentrations by County in January ')
 FL_winter_PM25_concen_fig
 
 #AQI .. same graph as above but on different time scales on Y .. so may be easier to read AQI ? not sure
 FL_winter_PM25_AQI <- FL_winter_PM25_egrid %>%
  ggplot(aes(x = Mean_Plant_NetGenMW, y = Average_Month_AQI_Value)) +
   geom_jitter() +
   #scale_x_log10() + #Heili suggested this!! better for outliers-can set ticks / break points on x axis too
  labs(x = 'Average Net Gen Capacity (MW)', y = 'Average PM25 AQI',
       title = 'January')
 FL_winter_PM25_AQI

```

## FLORIDA SUMMER
### Data Wrangling
```{r}
# data wrangling
#Florida 
summer_PM25_FL_data <-PM25_FL_data %>%
  filter(Month == 7) %>% #july
  group_by(County, County.FIPS.Code) %>%
  summarize(average_month_PM25_concentration = mean(Daily.Mean.PM2.5.Concentration),
            Average_Month_AQI_Value = mean(Daily.AQI.Value)) %>% #not sure which to do/difference rn
  ungroup()

#now combining the two data frames
FL_summer_PM25_egrid <- merge(egrid_data_FL_counties, summer_PM25_FL_data, by = 'County.FIPS.Code')
```

### Graphs
```{r}
#by mean PM25 concentration on y and Net Gen on x
FL_summer_PM25_concen <- FL_summer_PM25_egrid %>%
  ggplot(aes(x = Mean_Plant_NetGenMW, y = average_month_PM25_concentration)) +
   geom_jitter() +
  scale_x_log10() +
  labs(x = 'Mean Net Gen Capacities of Florida Counties (MW)', y = 'Average PM25 Concentrations by Counties (unit)',
       title = 'Florida Average Net Gen Capacities and \n PM25 Concentrations by County in July ')
 FL_summer_PM25_concen
 
 #AQI .. same graph as above but on different time scales on Y .. so may be easier to read AQI ? not sure
 FL_summer_PM25_AQI <- FL_summer_PM25_egrid %>%
  ggplot(aes(x = Mean_Plant_NetGenMW, y = Average_Month_AQI_Value)) +
   geom_jitter() +
   #scale_x_log10() +
  labs(x = 'Average Net Gen Capacity (MW)', y = 'Average PM25 AQI',
       title = 'July ')
 FL_summer_PM25_AQI
```
```{r}
#FLORIDA GRAPHS
# cool patchwork feature Heili told me about!! wow

FL_winter_PM25_AQI + FL_summer_PM25_AQI +
    plot_layout(axes = 'collect') + #to consolidate x axis titles 
  plot_annotation(tag_levels = 'A') +
  plot_annotation('2022 Average Net Gen Capacities and PM25 AQI \n for Florida Counties in January and July', 
                  caption = 'These two plots show the combined average Net Gen capacities for coal-burning power plants \n averaged by Florida counties compared with the average EPA PM25 Air Quality Index \n of the specified month in those counties. Plot A shows the average PM25 AQI \n in January and Plot B shows AQI in July. ') 
  

```


```{r}
#winter continued ...

#Georgia
winter_PM25_GA_data<-PM25_GA_data

#Mississippi 
winter_PM25_MS_data<-PM25_MS_data

#North Carolina
winter_PM25_NC_data<-PM25_NC_data

#South Carolina
winter_PM25_SC_data<-PM25_SC_data

#Tennessee
winter_PM25_TN_data<-PM25_TN_data

#Alabama
winter_PM25_AL_data<-PM25_AL_data

```

# Summer Data (July)




```{r}
#Georgia
summer_PM25_GA_data<-PM25_GA_data

#Mississippi 
summer_PM25_MS_data<-PM25_MS_data

#North Carolina
summer_PM25_NC_data<-PM25_NC_data

#South Carolina
summer_PM25_SC_data<-PM25_SC_data

#Tennesse
summer_PM25_TN_data<-PM25_TN_data

#Alabama
summer_PM25_AL_data<-PM25_AL_data

```

# Summer - Combined chart of ALL counties

```{r}
#Data wrangling for PM25

summer_PM25_data_all <- PM25_data_all %>%
  filter(Month == 7) %>% #july
  group_by(County, County.FIPS.Code) %>%
  summarize(average_month_PM25_concentration = mean(Daily.Mean.PM2.5.Concentration),
            Average_Month_AQI_Value = mean(Daily.AQI.Value)) %>% #not sure which to do/difference rn
  ungroup()

#Data wrangling for egrid data
egrid_data_counties_all <- egrid_data %>%
  relocate(Plant.FIPS.county.code, 
           Plant.county.name, `Plant.annual.net.generation.(MWh)`) %>% #so i can see what im working with
  select(Plant.FIPS.county.code, 
         Plant.county.name,
         `Plant.annual.net.generation.(MWh)`, `Plant.state.abbreviation`) %>% #only important ones i think
  rename(County.FIPS.Code = Plant.FIPS.county.code) %>% #to merge egrid & PM25 data frames by common col
  group_by(County.FIPS.Code, Plant.county.name, Plant.state.abbreviation) %>% #change group by to new column name
  summarize(Mean_Plant_NetGenMW = mean(`Plant.annual.net.generation.(MWh)`)) %>%
  ungroup()

#now combining the two data frames
summer_PM25_egrid_all <- merge(egrid_data_counties_all, summer_PM25_data_all, by = 'County.FIPS.Code')
write.csv(summer_PM25_egrid_all, file = "summer_PM25_egrid_all.csv", row.names = FALSE)
```

```{r}
#graphs for combined data

#by mean PM25 concentration on y and Net Gen on x
all_summer_PM25_concen_fig <- summer_PM25_egrid_all %>%
  ggplot(aes(x = Mean_Plant_NetGenMW, y = average_month_PM25_concentration)) +
   geom_jitter() +
  scale_x_log10() +
  labs(x = 'Mean Net Gen Capacities of all Counties (MW)', y = 'Average PM25 Concentrations by County (unit)',
       title = 'Average Net Gen Capacities and \n PM25 Concentrations by County in July ')
all_summer_PM25_concen_fig
 
 #AQI .. same graph as above but on different time scales on Y .. so may be easier to read AQI ? not sure
 all_summer_PM25_AQI <- summer_PM25_egrid_all %>%
  ggplot(aes(x = Mean_Plant_NetGenMW, y = Average_Month_AQI_Value)) +
   geom_jitter() +
   scale_x_log10() + #Heili suggested this!! better for outliers-can set ticks / break points on x axis too
  labs(x = 'Average County Net Gen Capacity (MW)', y = 'Average PM25 AQI',
       title = 'July ')
 all_summer_PM25_AQI
```


# Winter - Combined chart of ALL counties
```{r}
#Data wrangling for PM25

winter_PM25_data_all <- PM25_data_all %>%
  filter(Month == 1) %>% #january
  group_by(County, County.FIPS.Code) %>%
  summarize(average_month_PM25_concentration = mean(Daily.Mean.PM2.5.Concentration),
            Average_Month_AQI_Value = mean(Daily.AQI.Value)) %>% #not sure which to do/difference rn
  ungroup()


#now combining the two data frames
winter_PM25_egrid_all <- merge(egrid_data_counties_all, winter_PM25_data_all, by = 'County.FIPS.Code')
write.csv(winter_PM25_egrid_all, file = "winter_PM25_egrid_all.csv", row.names = FALSE)
```

```{r}
#graphs for combined data

#by mean PM25 concentration on y and Net Gen on x
all_winter_PM25_concen_fig <- winter_PM25_egrid_all %>%
  ggplot(aes(x = Mean_Plant_NetGenMW, y = average_month_PM25_concentration)) +
   geom_jitter() +
  scale_x_log10() +
  labs(x = 'Mean Net Gen Capacities of all Counties (MW)', y = 'Average PM25 Concentrations by County (unit)',
       title = 'Average Net Gen Capacities and \n PM25 Concentrations by County in January ')
all_winter_PM25_concen_fig
 
 #AQI .. same graph as above but on different time scales on Y .. so may be easier to read AQI ? not sure
 all_winter_PM25_AQI <- winter_PM25_egrid_all %>%
  ggplot(aes(x = Mean_Plant_NetGenMW, y = Average_Month_AQI_Value)) +
   geom_jitter() +
   scale_x_log10() + #Heili suggested this!! better for outliers-can set ticks / break points on x axis too
  labs(x = 'Average County Net Gen Capacity (MW)', y = 'Average PM25 AQI',
       title = 'January ')
 all_winter_PM25_AQI
```

```{r}
#All counties GRAPHS

all_winter_PM25_AQI + all_summer_PM25_AQI + 
  plot_layout(axes = 'collect') + #to consolidate x axis titles 
  plot_annotation(tag_levels = 'A') +
  plot_annotation('2022 Average Net Gen Capacities and PM25 AQI \n for All Counties in January and July', 
                  caption = 'These two plots show the combined average Net Gen capacities for coal-burning power plants \n averaged by counties in 7 Southeastern states compared with the average EPA PM25 Air Quality Index \n of the specified month. Plot A shows the average PM25 AQI in January and Plot B shows AQI in July. \n 
                  *These graphs do use scale_x_log10 for the X-Axis. ') 
  

#different y ranges which visually isnt great - can change later

```
