---
title: "tidycensus_data"
output: pdf_document
date: "2025-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(rvest)
library(dataRetrieval)
library(tidycensus)
library(tidyverse);library(lubridate);library(viridis);library(here)
here()
```

### Using `tidycensus` to extract census data
The `tidycensus` package uses the Census' API to pull data into our coding environment. To use it, however, you first need to get a free Census API key at <https://api.census.gov/data/key_signup.html>. After that, it's pretty easy to fetch data into your coding environment. 

We'll use it to extract total population of counties in NC. 

You'll also need to look up the IDs of the variables you want to extract. Those are found here:
<https://walker-data.com/tidycensus/articles/basic-usage.html#searching-for-variables>


```{r}
# population data by county for all the states

#Apply the API key
census_api_key("d4ad7874c996c7a83fec0a81466309ddc07d8e22a", install = TRUE)

v17 <- load_variables(2022, 'acs5', cache = TRUE)
v17

AL_pop <- 
  get_acs(geography = "county",
          state = "AL",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

FL_pop <- 
  get_acs(geography = "county",
          state = "FL",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

GA_pop <- 
  get_acs(geography = "county",
          state = "GA",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

MS_pop <- 
  get_acs(geography = "county",
          state = "MS",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

NC_pop <- 
  get_acs(geography = "county",
          state = "NC",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

SC_pop <- 
  get_acs(geography = "county",
          state = "SC",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

TN_pop <- 
  get_acs(geography = "county",
          state = "TN",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

```


```{r}
#find another variable 


```

