---
title: "tidycensus_data"
output: pdf_document
date: "2025-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(rvest)
library(dataRetrieval)
library(tidycensus)
library(tidyverse);library(lubridate);library(viridis);library(here)
here()
```

### Using `tidycensus` to extract census data
The `tidycensus` package uses the Census' API to pull data into our coding environment. To use it, however, you first need to get a free Census API key at <https://api.census.gov/data/key_signup.html>. After that, it's pretty easy to fetch data into your coding environment. 

We'll use it to extract total population of counties in NC. 

You'll also need to look up the IDs of the variables you want to extract. Those are found here:
<https://walker-data.com/tidycensus/articles/basic-usage.html#searching-for-variables>


```{r}
# population data by county for all the states

#Apply the API key
census_api_key("d4ad7874c996c7a83fec0a81466309ddc07d8e22a", install = TRUE, overwrite = TRUE)

v17 <- load_variables(2022, 'acs5', cache = TRUE)
v17

AL_pop <- 
  get_acs(geography = "county",
          state = "AL",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

FL_pop <- 
  get_acs(geography = "county",
          state = "FL",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

GA_pop <- 
  get_acs(geography = "county",
          state = "GA",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

MS_pop <- 
  get_acs(geography = "county",
          state = "MS",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

NC_pop <- 
  get_acs(geography = "county",
          state = "NC",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

SC_pop <- 
  get_acs(geography = "county",
          state = "SC",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

TN_pop <- 
  get_acs(geography = "county",
          state = "TN",
          variables = 'B01001_001',
          year = 2022) #gives back spatial features

```


```{r}
#median household income 

AL_median_income <- 
  get_acs(geography = "county",
          state = "AL",
          variables = 'B19013_001',
          year = 2022) #gives back spatial features

FL_median_income <- 
  get_acs(geography = "county",
          state = "FL",
          variables = 'B19013_001',
          year = 2022) #gives back spatial features

GA_median_income <- 
  get_acs(geography = "county",
          state = "GA",
          variables = 'B19013_001',
          year = 2022) #gives back spatial features

MS_median_income <- 
  get_acs(geography = "county",
          state = "MS",
          variables = 'B19013_001',
          year = 2022) #gives back spatial features

NC_median_income <- 
  get_acs(geography = "county",
          state = "NC",
          variables = 'B19013_001',
          year = 2022) #gives back spatial features

SC_median_income <- 
  get_acs(geography = "county",
          state = "SC",
          variables = 'B19013_001',
          year = 2022) #gives back spatial features

TN_median_income <- 
  get_acs(geography = "county",
          state = "TN",
          variables = 'B19013_001',
          year = 2022) #gives back spatial features
```
Housing & Transportation
Percentage of housing units with no vehicle available: B25044 (need to figure out how to work with this series)

```{r}
#Use B25044_001 (Total housing units) and B25044_003 (units with no vehicle) to figure out percent of households with no vehicle

AL_percent_no_vehicle <- 
  get_acs(geography = "county",
          state = "AL",
          variables = c('B25044_001', 'B25044_003'),
          year = 2022) %>% 
  select(GEOID, NAME, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate) %>%
  mutate(percent_no_vehicle = (`B25044_003` / `B25044_001`) * 100)

FL_percent_no_vehicle <- 
  get_acs(geography = "county",
          state = "FL",
          variables = c('B25044_001', 'B25044_003'),
          year = 2022) %>% 
  select(GEOID, NAME, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate) %>%
  mutate(percent_no_vehicle = (`B25044_003` / `B25044_001`) * 100)

GA_percent_no_vehicle <- 
  get_acs(geography = "county",
          state = "GA",
          variables = c('B25044_001', 'B25044_003'),
          year = 2022) %>% 
  select(GEOID, NAME, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate) %>%
  mutate(percent_no_vehicle = (`B25044_003` / `B25044_001`) * 100)

MS_percent_no_vehicle <- 
  get_acs(geography = "county",
          state = "MS",
          variables = c('B25044_001', 'B25044_003'),
          year = 2022) %>% 
  select(GEOID, NAME, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate) %>%
  mutate(percent_no_vehicle = (`B25044_003` / `B25044_001`) * 100)

NC_percent_no_vehicle <- 
  get_acs(geography = "county",
          state = "NC",
          variables = c('B25044_001', 'B25044_003'),
          year = 2022) %>% 
  select(GEOID, NAME, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate) %>%
  mutate(percent_no_vehicle = (`B25044_003` / `B25044_001`) * 100)

SC_percent_no_vehicle <- 
  get_acs(geography = "county",
          state = "SC",
          variables = c('B25044_001', 'B25044_003'),
          year = 2022) %>% 
  select(GEOID, NAME, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate) %>%
  mutate(percent_no_vehicle = (`B25044_003` / `B25044_001`) * 100)

TN_percent_no_vehicle <- 
  get_acs(geography = "county",
          state = "TN",
          variables = c('B25044_001', 'B25044_003'),
          year = 2022) %>% 
  select(GEOID, NAME, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate) %>%
  mutate(percent_no_vehicle = (`B25044_003` / `B25044_001`) * 100)

```
Housing cost burden: B25070 (also need to figure out how to work with this)

According to the U.S. Department of Housing and Urban Development (HUD), households spending more than 30% of their income on housing costs, including rent, mortgage payments, and utilities, are considered housing cost-burdened.

B25070_001: Total renter-occupied units
B25070_007 through B25070_010: Units spending 30% or more of income on rent
B25070_007: 30–34.9%
B25070_008: 35–39.9%
B25070_009: 40–49.9%
B25070_010: 50% or more

summing up variables 007 to 010 and dividing by total 001

```{r}

```

Race & Ethnicity
Race breakdown (White, Black, Asian, etc.): 
B02001_002: white
B02001_003: black
B02001_004: native
B02001_005: asian
B02001_006: pacific
B02001_007: other
B02001_008: two_or_more
^could potentially find percent of total pop of each? But can also scrap if too much work

```{r}

```